\documentclass[12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{tikz}
\makeatletter
\def\input@path{{../../style/}}
\makeatother

\usepackage{../../style/quiver}
\makeatletter
\def\input@path{{../../style/}}
\makeatother

\usepackage{../../style/scribe}

\begin{document}
Songyu Ye

\today

\hfill

\section{Introduction}
Let $G$ connected reductive group over $\C$ and $X$ smooth $G$
variety. There is an infinitesimal action of $G$ on $X$ given by the
Lie algebra $\mf g$ of $G$.
\begin{align*}
    \mf g \to \Gamma(X,T_X) \\
    U(\mf g) \to \Gamma(X,\cD_X)
\end{align*} and a localization functor \begin{align*}
    U(\mf g)-mod \to D(X)
\end{align*}
which takes a representation $V$ to the $D$ module $D(X)\otimes_{U(\mf g)} V$ on $X$.

\hfil 

In BB localization, $X$ is the flag variety and the localization functor is not a equivalence
but it becomes an equivalence once we restrict to those modules by which the center of $U(\mf g)$ 
acts trivially.

\section{Peter Weyl}

Recall the Peter-Weyl theorem: 

\begin{align*}
    \C[G] \cong \bigoplus_{\lambda} V_{\lambda} \otimes V_{\lambda}^*
\end{align*} where the sum is over all fd representations of $G$.
There is a filtration on this ring via dominance order: \begin{align*}
    \C[G]_{\leq \mu} = \bigoplus_{\lambda \leq \mu} V_{\lambda} \otimes V_{\lambda}^*
\end{align*} There is an algebra homomorphism \begin{align*}
    \mu: U(\mf g) \otimes U(\mf g) \to D_G
\end{align*} given by inclusion of left and right differential operators
so that the image of $\mu = U(\mf g) \otimes_{Z(U(\mf g))} U(\mf g)$. Also this map gives an induced filtration on 
$D_G$: \begin{align*}
    D_{G,\leq\mu} = \im\mu\cdot O(G)_{\leq\mu}
\end{align*}

\begin{example}
    $G = \SL_2$ so that we are $\C[a,b,c,d]/(ad-bc-1)$ and the filtration
    in question is by degree. Then we have the map \begin{align*}
        U(\mf sl_2) \otimes U(\mf sl_2) \to D_{\SL_2}
    \end{align*} given by \begin{align*}
        X\otimes 1 &\mapsto X \\
        1\otimes E &\mapsto -a^2E + c^2F + acH \quad \text{\red{check reference}}\\
        \vdots
    \end{align*}
\end{example}
and can check that the Casimir goes to the same place whether you put it on the left or right.

\section{Horocycle and Vinberg semigroup}

$B^-$ opposite borel $B\cap B^- = T$ torus. 
$N$ unipotent radical of $B$ and $N^-$ unipotent radical of $B^-$.
$Z$ center of $G$.
$G_ad$ adjoint group of $G$.
$T_ad = T/Z$.

\begin{definition}
    The horocycle space is \begin{align*}
        Y = G/N\times_T N^-\backslash G
    \end{align*} has an action of $G\times G \times T$ 
    and is a principle $T$ bundle over \begin{align*}
        G/B\times G/B^- 
    \end{align*}
\end{definition}
 The affine closure of $Y$ is $\Spec \gr O(G)$.

 \begin{proposition}
    [Vinberg] There is a smooth $G\times G\times T$ variety $V_\circ$
    which fits the following diagram. 
    % https://q.uiver.app/#q=WzAsNixbMSwwLCJWX1xcY2lyYyJdLFsyLDAsIkdcXHRpbWVzX3taKEcpfSBUIl0sWzIsMSwiXFxtYXRoYmJ7Q14qfV5yIl0sWzEsMSwiXFxtYXRoYmJ7QX1eciJdLFswLDEsIjAiXSxbMCwwLCJZIl0sWzAsM10sWzEsMF0sWzIsM10sWzEsMl0sWzQsM10sWzUsNF0sWzUsMF1d
\[\begin{tikzcd}
	Y & {V_\circ} & {G\times_{Z(G)} T} \\
	0 & {\mathbb{A}^r} & {\mathbb{C^*}^r}
	\arrow[from=1-1, to=1-2]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-2, to=2-2]
	\arrow[from=1-3, to=1-2]
	\arrow[from=1-3, to=2-3]
	\arrow[from=2-1, to=2-2]
	\arrow[from=2-3, to=2-2]
\end{tikzcd}\]
 \end{proposition} where $\pi$ is flat with smooth fibers, all maps are $G\times G\times T$ equivariant,
 where $G\times G$ acts on $\A^r$ trivially
and the action of $G\times G\times T$ on $V_\circ$ is free and 
the quotient by $T$ gives precisely the wonderful compactification.

\begin{example}
    Back to $\SL_2$: The horocycle space \begin{align*}
        Y = \C^2-0 \times \C^2-0 /\C^* = \text{rank $1$ matrices}
    \end{align*} has a map to $\P^1\times \P^1$ given by kernel and image.

    The diagram looks like % https://q.uiver.app/#q=WzAsNixbMCwwLCJZIl0sWzEsMCwiTV97MiwyfS0wIl0sWzEsMSwiQyJdLFswLDEsIjAiXSxbMiwwLCJHTF8yIl0sWzIsMSwiXFxtYXRoYmJ7Q31eKiJdLFswLDFdLFsxLDIsImRldCIsMl0sWzAsM10sWzMsMl0sWzQsMV0sWzUsMl0sWzQsNV1d
    \[\begin{tikzcd}
        Y & {M_{2,2}-0} & {GL_2} \\
        0 & C & {\mathbb{C}^*}
        \arrow[from=1-1, to=1-2]
        \arrow[from=1-1, to=2-1]
        \arrow["det"', from=1-2, to=2-2]
        \arrow[from=1-3, to=1-2]
        \arrow[from=1-3, to=2-3]
        \arrow[from=2-1, to=2-2]
        \arrow[from=2-3, to=2-2]
    \end{tikzcd}\]
\end{example}

\subsection{Construction}

\end{document}