
\documentclass[12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{mathtools}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{bookmark}


\makeatletter
\def\input@path{{../../style/}}
\makeatother

\usepackage{../../style/quiver}
\makeatletter
\def\input@path{{../../style/}}
\makeatother

\usepackage{../../style/scribe}
\usepackage{fancyhdr}

\usepackage{parskip}
\setlength{\parskip}{1em}
\setlength{\parindent}{0pt}

\DeclareMathOperator{\len}{len}
\newcommand{\fg}{\mathfrak g}
\newcommand{\Rees}{Rees}

\DeclareMathOperator{\Frac}{Frac}
\begin{document}

\lhead{Songyu Ye}
\rhead{\today}
\cfoot{\thepage}

\title{Title}
\author{Songyu Ye}
\date{\today}
\maketitle

\begin{abstract}
    Abstract
\end{abstract}

\tableofcontents

\section{The setup}
Recall that we have a line bundle $\mathcal L:=\mathcal L_{\det}$ on $\cX:=\cX_{G,g,I}$ and an invariant norm $\|\cdot\|$ on $X_*(T)_\mathbb R$.

For a point $x\in \cX(k)$ and a nontrivial $f:\Theta\to \cX$ with $f(1)\simeq x$, define
$$\mu(x,f)\;:=\;\frac{\mathrm{wt}_{\mathcal L}(f)}{\|\lambda_f\|},$$
where $\mathrm{wt}_{\mathcal L}(f)$ is the $\mathbb G_m$-weight on the fiber of $\mathcal L$ at the special point $f(0)$, and $\lambda_f$ is the associated cocharacter data of $f$ coming from the action of $\mathbb G_m$ on the object $f(0)$.

We wanted to prove the following theorem.
\begin{theorem}[Need to prove]
    For any $x\in \cX_{G,g,I}(k)$, there exists a (unique up to something) maximally destabilizing $\Theta$-filtration $f_{\mathrm{HN}}:\Theta\to \cX_{G,g,I}$ of $x$ with respect to $\mathcal L_{\det}$ and any invariant norm on $X_*(T)_\R$.
\end{theorem}

Constantin remarked that the uniqueness part of the theorem is not important for our application, and that we can get away with existence of a maximally destabilizing $\Theta$-filtration for each $x\in \cX(k)$, provided that we can organize the points $x$ into strata in such a way which mirrors the setup of the Kirwan Ness stratification.

\section{Kirwan Ness stratification}
This section follows \cite{kirwan}.
Suppose we have linear action of a reductive group $G$
on a projective variety $X$, singular or nonsingular, defined over an algebraically closed field. Let $T \subset G$ be a maximal torus, $V$ the representation giving the linearization.
Pick an invariant norm on the cocharacter lattice of $G$. Then Kirwan and Ness construct a stratification of $X$ into locally closed subvarieties \begin{align*}
    X = \bigsqcup_{\beta \in \mathcal B} S_\beta,
\end{align*} \textbf{The index set $\mathcal B$ is in correspondence with connected components $Z \subset X^{\beta}$ of the fixed locus of a dominant cocharacter $\beta$ of $T$ such that the semistable locus $Z^\circ$ of the divided action of the Levi subgroup $L_\beta$ on $Z$ is nonempty.}

If the fixed loci is connected, then the index set $\mathcal B$ can be identified with the set of
$G$ conjugacy classes of rays in $X_*(G) \otimes \R$ together with the condition that the associated “center” $Z_\beta^{ss}$ is nonempty. Note that every cocharacter of $G$ is $G$-conjugate to a dominant cocharacter of $T$. It is harmless to choose a rational dominant cocharacter $\beta$ in each conjugacy class as they determine the same parabolic subgroup $P_\beta$ and the same stratum $S_\beta$. Note that one cannot simply group the components together under a single conjugacy class of rays, as each connected component $Z_i$ could possibly a different weight of the linearized line bundle.


Then we have a weight decomposition \begin{align*}
    V = \bigoplus_{\chi} V_\chi,
\end{align*}
and so for any point $x\in X$, we can write its homogeneous coordinates as $x = [v]$ for some $v\in V$ with $v = \sum_\chi v_\chi$. For a point $x$, let $W_x$ be the set of weights appearing in its support, meaning the set of $\chi$ such that $v_\chi \neq 0$. Then Kirwan identifies $\beta$ as the closest point to $0$ in $\mathrm{Conv}(W_x)$.

Alternatively, we can identify $\beta$ as the $G$-conjugacy class of ray which minimzes the following function on $X_*(G) \otimes \R$:
\begin{align*}
    \lambda \mapsto \frac{\mu(x,\lambda)}{\|\lambda\|}.
\end{align*}
where $\mu(x,\lambda)$ is defined as the minimum $\lambda$-weight of the nonzero coordinates of $x$.
\begin{align*}
\mu(x,\lambda)= \min_{\chi \in W_x}\langle \chi, \lambda \rangle
\end{align*}

\begin{proposition}
    [Hilbert-Mumford criterion for semistability] A point $x$ is semistable if and only if $\mu(x,\beta) \geq 0$ for all dominant cocharacters $\beta$.
\end{proposition}

For each $\beta\in \mathcal B$, pick a dominant rational representative of the corresponding conjugacy class of rays which we also denote by $\beta$. The unstable strata are indexed by those $Z$ with dominant $\beta$ for which
the semistable locus $Z^\circ \subset Z$ of the divided $L$-action on $\mathcal L$
is not empty. 

In particular we look at the fixed locus 
$X^\beta = \{ x \in X \mid \beta(t)\cdot x = x \ \forall t \}$ and take a connected component $Z \subset X^\beta$. Now define the attracting set:
\[Y=\{ x \in X \mid \lim_{t\to\infty} \lambda(t)\cdot x \in Z \}\] which gives us a natural map $\varphi: Y \to Z$ defined by the limit of the $T$-flow. 

\begin{proposition}
    $\varphi : Y \to Z$ is a locally trivial fibration in affine spaces. At a point $z\in Z$, the tangent space decomposes into weight spaces under $\beta$ as:
\begin{align*}
    T_zX = T_zZ \oplus (T_zX)_{>0}
\end{align*}
The positive-weight directions integrate to affine fibers.
\end{proposition}

Now we describe what $Z^\circ$ is. Let $L = Z_G(\beta)$
be the Levi subgroup. $\beta$ acts on the fiber of the linearized line bundle $\mathcal L$ over $Z$ by the Hilbert-Mumford weight $\mu(x,\beta)$. Let $\chi$ be the character of $L$ given by this action. Twisting down by $\chi$, the action of $L$ on $Z$ is trivial on the fiber of the linearized line bundle, and so we can talk about semistability for the $L$-action on $Z$ with respect to the trivial linearization. We define $Z^\circ$ to be the semistable locus for this GIT problem, and put $Y^\circ = \varphi^{-1}(Z^\circ)$. Note that if we did not twist down by $\chi$, then the action on the fiber would be nontrivial and so the semistable locus would be empty, since $\beta: \mathbb G_m \to L$ is a one-parameter subgroup of $L$ which violates the Hilbert-Mumford criterion for semistability.

To recover the stratum $S_\beta$, we take the $G$-orbit of $Y^\circ$, i.e. $S = G \cdot Y^\circ$ and then one can show that we have the following isomorphism of $G$-varieties:
\[S \cong G \times^{P_{\beta}} Y^\circ\]
where we are dividing by the relation \begin{align*}
    (g, y) \sim (gp^{-1}, py) \quad \forall p \in P_\beta.
\end{align*}
\begin{proposition}
In this setup, we have the following properties:
\begin{enumerate}
\item[(i)][Bialynicki-Birula decomposition]
$Y$ is a fiber bundle over $Z$, with affine spaces as fibers, under the
morphism $\varphi$ defined by the limiting value of the $T$-flow. The fibers of the affine bundle are the positive weight spaces in the tangent space of $X$ at the limit point.

\item[(ii)]
$Y$ is stabilized by the parabolic subgroup $P \subset G$ whose nilpotent
Lie algebra radical $\mathfrak u$ is spanned by the negative $T$-eigenspaces
in $\mathfrak g$.

\item[(iii)]
$S$ is equivariantly isomorphic to $G \times^P Y^\circ$. In particular \begin{align*}
    S \cong G \times^P Y^\circ \cong (G \times^P Z^\circ) \times \mathbb A^n 
\end{align*} where the fibers of the affine bundle are the positive weight spaces in the tangent space of $X$ at the limit point, if we let $P$ act on $Z^\circ$ via its reductive quotient $L$.

\item[(iv)]
The various $S$, together with $X^\circ = X^{ss}$, smoothly stratify $X$.

\item[(v)]
$Z^\circ$ has a projective, good quotient under $L$; $X^{ss}$ has a good projective quotient under $G$.
\end{enumerate}
\end{proposition}

\section{Lattices in $\SL(2)$}
Let $F=\C((z))$, $\mathcal O=\C[[z]]$, $G=\SL_2(F)$.
Let
$$\Lambda_0:=\mathcal O e_1\oplus \mathcal O e_2,\qquad
\Lambda_1:=\mathcal O e_1\oplus z\,\mathcal O e_2,$$
and
$$P_0:=\Stab_G(\Lambda_0)=\SL_2(\mathcal O),\qquad
P_1:=\Stab_G(\Lambda_1).$$

Let
\[
S:=\begin{pmatrix}1&0\\0&z\end{pmatrix}\in \GL_2(F).
\]
Then $S\Lambda_0=\Lambda_1$. Hence
\[
P_1=\Stab_G(\Lambda_1)=\Stab_G(S\Lambda_0)=S\,\Stab_G(\Lambda_0)\,S^{-1}=S P_0 S^{-1},
\]
because conjugation by $S$ preserves determinant 1, so $S P_0 S^{-1}\subset \SL_2(F)$ and it is exactly the stabilizer of $S\Lambda_0$.

Conjugation by $S$ gives a bijection of double coset sets
\begin{align*}
    P_1\backslash G/P_1 &\xrightarrow{\sim} P_0\backslash G/P_0,\\
    P_1 g P_1 &\mapsto P_0 (S^{-1} g S) P_0.
\end{align*}

Moreover we know that $P_0\backslash G/P_0$ is in bijection with the set of dominant cocharacters of $\SL_2$, which is $\Z_{\geq 0}$, via the Cartan decomposition \begin{align*}
    G = \bigsqcup_{n\geq 0} P_0 \begin{pmatrix} z^n & 0 \\ 0 & z^{-n} \end{pmatrix} P_0.
\end{align*}

Hence $P_1\backslash G/P_1$ is also in bijection with $\Z_{\geq 0}$.

\subsection{Heinloth's weight formula for $\SL_2$ and the nodal contribution on $X_{\mathrm{Solis}}$}

Let $C$ be a smooth projective curve and $E$ a principal $\SL_2$--bundle on $C$.
Let $\mathcal L_{\det}$ denote the determinant of cohomology line bundle on
$\Bun_{\SL_2}(C)$ associated to the adjoint representation $\Ad$.

\subsubsection*{1. Smooth case: Heinloth's formula for $\SL_2$}

A maximal parabolic of $\SL_2$ is a Borel subgroup $B\subset \SL_2$, so a reduction
$E_B$ is equivalently a line subbundle
\[
L\ \subset\ V:=E\times^{\SL_2}\C^2
\qquad\text{with}\qquad
\det(V)\cong\mathcal O_C.
\]
Let $T\subset B$ be a maximal torus.  As a $T$--representation,
\[
\mf{sl}_2 \;=\; \mf t \oplus \mf g_+ \oplus \mf g_-,
\]
where $\mf g_+$ (resp.\ $\mf g_-$) is the positive (resp.\ negative) root space
and $\mf t$ has weight $0$.
Under the Borel reduction determined by $L$, the positive root bundle is
\[
E_B\times^B \mf g_+ \ \cong\ L^{\otimes 2}.
\]

Let $f:\Theta=[\A^1/\Gm]\to \Bun_{\SL_2}(C)$ be the very close degeneration
$\Rees(E_B,\check\omega)$ attached to the fundamental cocharacter $\check\omega$
(the unique maximal parabolic direction for $\SL_2$).
Heinloth's computation (\cite[\S1.F.c]{Heinloth}) gives
\begin{equation}
\label{eq:heinloth-sl2-weight}
\wt_{\mathcal L_{\det}}(f)
\;=\;
2\sum_{i>0} i\cdot \deg\!\big(E_{0,\check\omega}\times^{L_{\check\omega}}\Lie(G)_i\big)
\;=\;
2\deg(L^{\otimes 2})
\;=\;
4\deg(L),
\end{equation}
up to the choice of convention $t\to 0$ versus $t\to\infty$ in the definition of the weight.

\subsubsection*{2. The nodal/bubble contribution on $X_{\mathrm{Solis}}$}

Now let $x\in D_0\subset X_{\mathrm{Solis}}$ be a hyperspecial boundary point.
Thus $x$ is represented by an admissible $\SL_2$--bundle $E'$ on a twisted
modification
\[
m:C'\longrightarrow C_0
\]
of the nodal curve $C_0$, obtained by inserting an exceptional chain over the node.
Let $f:\Theta\to X_{\mathrm{Solis}}$ be a very close degeneration with $f(1)=x$ and
special fiber $f(0)$ an associated graded object carrying a Borel reduction $E'_B$.
Let
\[
L'\ \subset\ V':=E'\times^{\SL_2}\C^2
\]
be the corresponding line subbundle on $C'$ (so $\det(V')\cong\mathcal O_{C'}$).
The same representation--theoretic identification as above gives
\[
E'_B\times^B \mf g_+\ \cong\ (L')^{\otimes 2}.
\]
Therefore Heinloth's degree formula applies verbatim, but with degree computed on $C'$:
\begin{equation}
\label{eq:solis-weight-total-degree}
\wt_{\mathcal L_{\det}}(f)
\;=\;
2\,\deg_{C'}\!\big((L')^{\otimes 2}\big)
\;=\;
4\,\deg_{C'}(L').
\end{equation}

Write $C'$ as the union of the strict transform of the normalization and the
exceptional chain:
\[
C' \;=\; \widetilde C \ \cup\  (\text{exceptional chain}),
\]
with intersection along the two attaching points.  Additivity of degree gives
\begin{equation}
\label{eq:degree-split}
\deg_{C'}(L')
\;=\;
\deg_{\widetilde C}\!\big(L'|_{\widetilde C}\big)
\;+\;
\deg_{\mathrm{chain}}\!\big(L'|_{\mathrm{chain}}\big).
\end{equation}
The first term is the usual Harder--Narasimhan/Rees contribution coming from the
smooth part.  The second term is the nodal/bubble contribution.

In the hyperspecial stratum $D_0$, the local gluing class at the node is labeled
by a dominant coweight
\[
\lambda\in X_*(T)_+\cong\Z_{\ge 0}.
\]
Geometrically, the associated graded local model is a bubble carrying the split
bundle
\[
\mathcal O_{\P^1}(\lambda)\oplus \mathcal O_{\P^1}(-\lambda),
\]
and the graded Borel reduction picks out the summand $\mathcal O_{\P^1}(\lambda)$.
Hence on each bubble component we have
\[
L'|_{\P^1}\cong \mathcal O_{\P^1}(\lambda),
\qquad\text{so}\qquad
\deg_{\P^1}(L'|_{\P^1})=\lambda.
\]
For an exceptional chain, the same statement holds with $\lambda$ equal to the
\emph{total} relative position across the chain (distributed among components), so
\begin{equation}
\label{eq:node-degree}
\deg_{\mathrm{chain}}\!\big(L'|_{\mathrm{chain}}\big)=\lambda.
\end{equation}

Combining \eqref{eq:solis-weight-total-degree}, \eqref{eq:degree-split}, and
\eqref{eq:node-degree} yields the advertised decomposition of the weight:
\begin{equation}
\label{eq:solis-weight-final}
\wt_{\mathcal L_{\det}}(f)
\;=\;
4\Bigl(\deg_{\widetilde C}\!\big(L'|_{\widetilde C}\big)\;+\;\lambda\Bigr),
\qquad (\SL_2,\ x\in D_0).
\end{equation}

The $\Theta$-degeneration determines a B-reduction $E'_B$ of the graded object $E'$ on $C'$, $L'$ is the line bundle associated to that B-reduction (equivalently the line subbundle $L'\subset V'$), and the nodal contribution is $\deg(L'|_{\text{chain}})$, which can vary with the reduction even when the ambient hyperspecial double-coset label $\lambda$ is fixed.

\subsubsection*{A. Degree on a nodal curve: normalization + a len correction}

Let $C_0$ be an integral projective nodal curve with a single node $p\in C_0$,
and let $\nu:\widetilde C\to C_0$ be the normalization. Write
$\nu^{-1}(p)=\{p_x,p_y\}$.

\begin{lemma}[Degree of a rank-$1$ torsion-free sheaf at a node]
\label{lem:degree-normalization-len}
Let $F$ be a rank-$1$ torsion-free coherent sheaf on $C_0$. Then
$\widetilde F:=\nu^*F$ is a line bundle on $\widetilde C$, and there is a
canonical exact sequence
\begin{equation}
\label{eq:nu-push-exact}
0 \longrightarrow F \longrightarrow \nu_*\widetilde F \longrightarrow Q \longrightarrow 0,
\end{equation}
where $Q$ is a skyscraper sheaf supported at $p$. Moreover,
\begin{equation}
\label{eq:deg-split-len}
\deg_{C_0}(F)
\;=\;
\deg_{\widetilde C}(\widetilde F)
\;+\;
\len(Q).
\end{equation}
\end{lemma}

\begin{proof}
Since $F$ is rank-$1$ torsion-free on a nodal curve, it is locally free away
from $p$, and pulling back to the normalization removes the singularity, so
$\widetilde F=\nu^*F$ is a line bundle on $\widetilde C$. The natural adjunction
map $F\to \nu_*\nu^*F=\nu_*\widetilde F$ is an isomorphism away from $p$, hence
its cokernel $Q$ is supported at $p$, giving \eqref{eq:nu-push-exact}.

Taking Euler characteristics in \eqref{eq:nu-push-exact} yields
\[
\chi(C_0,F)=\chi(C_0,\nu_*\widetilde F)-\chi(C_0,Q)=\chi(\widetilde C,\widetilde F)-\len(Q),
\]
since $\chi(C_0,\nu_*\widetilde F)=\chi(\widetilde C,\widetilde F)$ and
$\chi(C_0,Q)=\len(Q)$. The same argument with $F=\cO_{C_0}$ gives
\[
\chi(C_0,\cO_{C_0})=\chi(\widetilde C,\cO_{\widetilde C})-\len(\nu_*\cO_{\widetilde C}/\cO_{C_0}),
\]
and $\len(\nu_*\cO_{\widetilde C}/\cO_{C_0})=1$ for a node. Using the
definition $\deg_X(G):=\chi(X,G)-\chi(X,\cO_X)$ for rank-$1$ torsion-free $G$ on
a proper curve $X$, the len terms combine to give \eqref{eq:deg-split-len}.
(Equivalently, one may take \eqref{eq:deg-split-len} as the definition of
$\deg_{C_0}(F)$ for torsion-free rank-$1$ sheaves.)
\end{proof}

\subsubsection*{B. Interpreting the hyperspecial invariant $\lambda$ in the $D_0$ case}

In the hyperspecial stratum $D_0$ we do \emph{not} necessarily insert an
exceptional chain of $\P^1$'s. Instead the local invariant
\[
\lambda\in X_*(T)_+\cong \Z_{\ge 0}
\]
is read off from the hyperspecial double coset
$\SL_2(\cO)\backslash \SL_2(F)/\SL_2(\cO)$ via the Cartan decomposition:
\[
g \in \SL_2(\cO)\,
\begin{pmatrix} z^\lambda & 0 \\ 0 & z^{-\lambda}\end{pmatrix}
\,\SL_2(\cO).
\]

Let $f:\Theta\to X_{\mathrm{Solis}}$ be a very close degeneration with $f(1)=x$
and let $E'^{\mathrm{gr}}=f(0)$ be the associated graded object. The $\Theta$-limit
supplies a Borel reduction $E'^{\mathrm{gr}}_B$, hence a line subsheaf
\[
L \subset V:=E'^{\mathrm{gr}}\times^{\SL_2}\C^2
\]
of the associated rank-$2$ bundle on $C_0$. In general $L$ is only rank-$1$
torsion-free at the node. Applying Lemma~\ref{lem:degree-normalization-len}
to $F=L$ gives
\begin{equation}
\label{eq:D0-degree-normalization-plus-lambda}
\deg_{C_0}(L)
\;=\;
\deg_{\widetilde C}(\nu^*L)
\;+\;
\len(Q_L),
\qquad
Q_L:=\nu_*\nu^*L/L.
\end{equation}

\begin{remark}[Where $\lambda$ enters]
\label{rem:lambda-as-len}
For the \emph{associated graded} local model of type $\lambda$ (i.e.\ after
conjugating the gluing to the diagonal representative and taking the Borel line
picked out by the $\Theta$-limit), the failure of $L$ to be locally free at the
node is measured precisely by $\lambda$, in the sense that
\[
\len(Q_L)=\lambda.
\]
Thus in the $D_0$ case (with no exceptional chain inserted) the ``nodal term''
appearing in the weight computation is a \emph{len correction at the node},
not a degree on a bubble.
\end{remark}

\subsubsection*{C. Consequence for the $\mathcal L_{\det}$-weight in the $D_0$ case}

Heinloth's formula gives, for $\SL_2$ and the adjoint determinant line bundle,
\[
\wt_{\cL_{\det}}(f)=4\,\deg_{C_0}(L)
\]
(up to the $t\to 0$ versus $t\to\infty$ convention). Combining with
\eqref{eq:D0-degree-normalization-plus-lambda} and
Remark~\ref{rem:lambda-as-len} yields
\[
\wt_{\cL_{\det}}(f)
\;=\;
4\Bigl(\deg_{\widetilde C}(\nu^*L)\;+\;\lambda\Bigr),
\qquad (\SL_2,\ x\in D_0).
\]


\subsection{The $D_1$ case}
Changing the trivializations on each branch by parahoric gauge transformations $p_x\in \mathcal P_x$, $p_y\in \mathcal P_y$ changes $\gamma$ by
\[
\gamma \mapsto p_y^{-1}\,\gamma\,p_x.
\]
    Let $E_x$ be a $\mathcal{P}_x$-torsor on the $x$-disc and $E_y$ a $\mathcal{P}_y$-torsor on the $y$-disc.
    \begin{itemize}
        \item Over the punctured discs they become $G((z))$-torsors (since $\mathcal{P}_{\bullet} \hookrightarrow G((z))$).
        \item Choose trivializations on the punctured discs:
        \begin{align*}
            \tau_x: E_x|_{D_x^\times} &\xrightarrow{\sim} G((z)),\\
            \tau_y: E_y|_{D_y^\times} &\xrightarrow{\sim} G((z)).
        \end{align*}
        \item The gluing isomorphism (descent datum) is an isomorphism of $G((z))$-torsors
        \begin{align*}
            \phi: E_x|_{D_x^\times} &\xrightarrow{\sim} E_y|_{D_y^\times}.
        \end{align*}
    \end{itemize}

Let us explain how to model descent datum on the nodal curve. 
Let $\cP= \cL\ltimes \cU$ be the parahoric with Levi $\cL$ and pro-unipotent radical $\cU$. Solis sets
\[
P^\Delta := \Delta(\cL) \ltimes (\cU\times \cU) \subset LG\times LG\]
Consider the map
\[
\Phi: LG\times LG \longrightarrow LG, \qquad (g_1,g_2)\longmapsto g_2\,g_1^{-1}.
\]

If $(g_1,g_2)$ is changed by an element of $P^\Delta$, say
\[
(g_1,g_2)\mapsto (g_1 u_1 \ell,\ g_2 u_2 \ell)
\quad\text{with }u_i\in U,\ \ell\in L,
\]
then
\[
g_2 u_2 \ell \cdot (g_1 u_1 \ell)^{-1}
=
g_2\ (u_2 \,\ell\,\ell^{-1}\,u_1^{-1})\ g_1^{-1}
=
g_2\,u_2\,u_1^{-1}\,g_1^{-1}.
\]
So the Levi cancels and $\Phi$ changes by
\[
\Phi \mapsto u_2\cdot \Phi \cdot u_1^{-1}.
\]

Conclusion: $\Phi$ descends to a map of quotients
\[
(LG\times LG)/\cP^\Delta \longrightarrow U\backslash LG/U.
\]
This quotient is the framed version of the local moduli at the node:
\begin{enumerate}
    \item $g_1$ is a trivialization on the $u$-branch punctured disc,
    \item $g_2$ is a trivialization on the $v$-branch punctured disc,
    \item modding out by $\cU\times\cU$ is change trivializations by parahoric-unipotent gauge on each branch.
    \item modding out by $\Delta(\cL)$ forces the Levi parts on the two branches to match (this is the descent compatibility across the node).
\end{enumerate}



We can further quotient
\[
U\backslash LG/U \twoheadrightarrow \cP\backslash LG/\cP
\]
because $\cP=L\ltimes U$ and this recovers the relative position invariants $\cP\backslash LG/\cP \cong W_I\backslash \widetilde{W}/W_I$ from which we will extract the nodal contribution to the weight.

The local parameter space before taking invariants is refined
\[(LG\times LG)/P^\Delta \to U\backslash LG/U \to \cP\backslash LG/\cP\]

The local node chart $(LG\times LG)/\mathcal{P}^\Delta$ refines the relative-position in $\mathcal{P}\backslash LG/\mathcal{P}\simeq W_I\backslash\widetilde{W}/W_I$. A $\Theta$-filtration supplies a $1$-PS $\lambda:\mathbb{G}_m\to \Delta(L)$ which acts with strictly positive weights on the pro-unipotent factor $U$. Therefore the $\mathbb{G}_m$-limit $t\to 0$ contracts the $U$-coordinates in the corresponding Bruhat/Schubert cell, and the limit lands in the unique $T$-fixed point in the Schubert closure, i.e. the point indexed by the minimal-len double-coset representative $w_{\min}\in W_I\backslash\widetilde{W}/W_I$.

This aids our calculation of the Hilbert Mumford weight as follows. Given a $\Theta$-degeneration $f:\Theta\to X_{\mathrm{Solis}}$. Over $\A^1$ you get a family of line bundles (your $L'$ coming from the reduction, or the positive-root line $(L')^{\otimes 2}$ in $\ad$). The degree on the exceptional chain is constant in flat families, so you may compute it on the special fiber $t=0$.

So you are allowed to replace your original local point (with complicated $U$-coordinates) by its $\Gm$-limit (the associated graded) without changing the degree you are computing.

Locally at the node, your object gives a point in a Schubert stratum of the partial affine flag variety $\Fl_I = LG/\mathcal{P}$ (or a refinement thereof). Flowing by the cocharacter $\lambda$ contracts $U$ and lands at some $T$-fixed point in the Schubert closure. The $T$-fixed points in $LG/\mathcal{P}$, under the action of $T$ on $LG/\mathcal{P}$ via left multiplication, are precisely $\{w\mathcal{P}\}_{w \in \widetilde{W}}$, where $w\mathcal{P} \in LG/\mathcal{P}$.


I suspect that the point indexed by the minimal double-coset representative
$w_{\min}\in W_I\backslash \widetilde W/W_I$.

Thus the local geometry of your degeneration becomes "the $T$-fixed point $w_{\min}$" plus a chain of $\P^1$'s corresponding to the simple reflections appearing in a reduced word for $w_{\min}$.

On $\Fl_I$, the Picard group is generated by the standard line bundles $\mathcal L_\alpha$ whose degrees are normalized by
\[
\deg(\mathcal L_\alpha|_{\P^1_\alpha})=1
\]
on the basic Schubert curve $\P^1_\alpha$ corresponding to a simple affine reflection $\alpha\notin I$.

So if your local line bundle (coming from the fundamental weight/Borel reduction) is the pullback of one of these generators (or a known combination), then its degree on any $1$–cycle is computed by counting how many times that $1$–cycle traverses the corresponding $\P^1_\alpha$'s. Take a reduced expression
$w_{\min}=s_{\alpha_1}s_{\alpha_2}\cdots s_{\alpha_r}$
in simple affine reflections (with $\alpha_i\notin I$ in the partial flag setting). The corresponding Schubert variety contains a canonical chain of Schubert curves
\[
\P^1_{\alpha_1}\cup \P^1_{\alpha_2}\cup\cdots\cup \P^1_{\alpha_r}
\]
connecting the base point to $w_{\min}$.

Then for a line bundle $\mathcal L$ whose degree on $\P^1_\alpha$ is known, you get
\[
\deg(\mathcal L\text{ on that chain})
=
\sum_{j=1}^r \deg\bigl(\mathcal L|_{\P^1_{\alpha_j}}\bigr).
\]

In $\SL_2$, $D_1$ case there is only one surviving simple direction (say $\alpha_0$), so every reduced word for $w_{\min}$ uses $\alpha_0$ exactly $n$ times, and
\[
\deg(\mathcal L|_{\text{chain}})=n\cdot \deg(\mathcal L|_{\P^1_{\alpha_0}}).
\]
With the standard normalization $\deg(\mathcal L|_{\P^1_{\alpha_0}})=1$, this is just $n$.

That is exactly why the "combinatorial number" attached to $w_{\min}$ becomes the degree.

\noindent\rule{\textwidth}{0.5pt}

How this feeds your $\mathcal L_{\det}$-weight calculation

Heinloth's weight for $\SL_2$ is
\[
\wt_{\mathcal L_{\det}}(f)=2\deg\big(E'_B\times^B\mf g_+\big)
=2\deg\big((L')^{\otimes 2}\big)=4\deg(L').
\]
So once you identify $\deg(L'|_{\text{chain}})$ with the integer $n$ coming from $w_{\min}$, the nodal contribution to the weight is immediately $4n$ (up to your sign convention).

If you want the next step prove $\deg(L'|_{\text{chain}})=n$ the clean proof route is:
- identify the chain with the Schubert \(\P^1\)-chain in the partial affine flag variety attached to $w_{\min}$,
- use that \(\Pic(LG/P_1)\cong \Z\) and degrees are normalized on the basic Schubert \(\P^1\),
- and observe that $w_{\min}=(s_0s_1)^n$ traverses that \(\P^1\) exactly $n$ times.

On the node, after taking the associated graded, the local data gives a map from the exceptional chain (a chain of \(\P^1\)’s) into the relevant Schubert closure in the partial affine flag $LG/P_I$. Each irreducible component of the chain maps to a Schubert curve \(\P^1_\alpha\)

Recall we are trying to compute the degree of the determinant of cohomology line bundle associated to the adjoint representation.

\section{References}
Teleman Quantization Conjecture Revisited.
Kirwan Cohomology of quotients in symplectic and algebraic geometry, 1984.
\end{document}